---
title: "HW5"
author: "Mason Bosley"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(rpart.plot)
library(tidymodels) 
tidymodels_prefer() # Resolves conflicts, prefers tidymodel functions
set.seed(123)

nba2021per36 <- read.csv("Per36-Player.csv")


nba2021_2022salaries <- read.csv("2021-2022-Player-Salaries.csv")
names(nba2021_2022salaries)[1] <- "Player"

nba2021per36 <- nba2021per36 %>% left_join(nba2021_2022salaries,by=c('Player')) %>%
  select(-AllStar)  %>%
  distinct(Player, .keep_all = TRUE) %>%
  filter(!is.na(Salary)) %>%
  filter(MP > 300) %>%
  na.omit()
head(nba2021per36)

```

```{r}
ggplot(nba2021per36, aes(x=`TRB`, y= `AST`)) +
    geom_point()
```

**The goal of the following code/models is to group players into the 5 position groups (Point Guard, Shooting Guard, Small Forward, Power Forward, Center) using K-means clustering. We have selected statistics that differentiate smaller players from larger players to make position predictions.**
```{r}
nba2021per36_sub <- nba2021per36 %>%
    select(STL, BLK, AST, TRB, FG., FT., X3PA)

# Run k-means for k = centers = 3
set.seed(253)
kclust_k5 <- kmeans(nba2021per36_sub, centers = 5)

# Display the cluter assignments
kclust_k5$cluster


nba2021per36 <- nba2021per36 %>%
    mutate(kclust_5 = factor(kclust_k5$cluster))
```

```{r}
ggplot(nba2021per36, aes(x=`BLK`, y= `AST`, color=kclust_5)) +
    geom_point()
```

**The following scaled graph uses blocks (a stat more common among large players) and assists (smaller players). It seems to at least differentiate between point guards (high assists, low blocks, shown in green) and centers (low assists, high blocks, shown in red). It is difficult to account for the data points in the lower left of the graph, as unimpactful players could be at any position.**
```{r}
set.seed(253)
kclust_k5_scale <- kmeans(scale(nba2021per36_sub), centers = 5)
nba2021per36 <- nba2021per36 %>%
    mutate(kclust_5_scale = factor(kclust_k5_scale$cluster))

# Visualize the new cluster assignments
ggplot(nba2021per36, aes(x=`BLK`, y= `AST`, color=kclust_5_scale)) +
    geom_point()
```

```{r}
nba2021per36_sub <- nba2021per36 %>%
    select(STL, BLK, AST, TRB, FG., FT., X3PA)

# Run k-means for k = centers = 3
set.seed(253)
kclust_k3 <- kmeans(nba2021per36_sub, centers = 3)

# Display the cluter assignments
kclust_k3$cluster


nba2021per36 <- nba2021per36 %>%
    mutate(kclust_3 = factor(kclust_k3$cluster))

ggplot(nba2021per36, aes(x=`BLK`, y= `FG.`, color=kclust_3)) +
    geom_point()

set.seed(253)
kclust_k3_scale <- kmeans(scale(nba2021per36_sub), centers = 3)
nba2021per36 <- nba2021per36 %>%
    mutate(kclust_3_scale = factor(kclust_k3_scale$cluster))

# Visualize the new cluster assignments
ggplot(nba2021per36, aes(x=`BLK`, y= `FG.`, color=kclust_3_scale)) +
    geom_point()
```



Hierarchical Clustering

```{r}
# Random subsample of 50 penguins
set.seed(645)
#nba2021per36 <- nba2021per36 %>%
   # slice_sample(n = 50)

# Select the variables to be used in clustering
nba2021per36_sub <- nba2021per36 %>%
    select(BLK, AST, ORB)

# Summary statistics for the variables
summary(nba2021per36_sub)

# Compute a distance matrix on the scaled data
dist_mat_scaled <- dist(scale(nba2021per36_sub))

# The (scaled) distance matrix is the input to hclust()
# The method argument indicates the linkage type
hc_complete <- hclust(dist_mat_scaled, method = "complete")
hc_single <- hclust(dist_mat_scaled, method = "single")
hc_average <- hclust(dist_mat_scaled, method = "average")
hc_centroid <- hclust(dist_mat_scaled, method = "centroid")

# Plot dendrograms
plot(hc_complete)
plot(hc_single)
plot(hc_average)
plot(hc_centroid)
```
``` {r}
plot(hc_complete, labels = nba2021per36$Player)
```

```{r}
nba2021per36 <- nba2021per36 %>%
    mutate(
        hclust_height4 = factor(cutree(hc_complete, h = 6)), # Cut at height (h) 3
        hclust_num3 = factor(cutree(hc_complete, k = 5)) # Cut into 6 clusters (k)
    )

nba2021per36 %>%
    ggplot(aes(x=hclust_height4, color=AST,y=Pos, size=PF)) +
    geom_point()

nba2021per36 %>%
    ggplot(aes(x=hclust_num3, color=Pos,y=AST, size=ORB)) +
    geom_point()

library(rgl)
plot3d(nba2021per36$AST, nba2021per36$BLK, nba2021per36$ORB, col = nba2021per36$hclust_num3)
```

```{r}
nba2021pergame %>%
  ggplot(aes(x=Pos, y=Salary)) +
  geom_boxplot()

nba2021per36 %>%
  ggplot(aes(x=Pos, y=Salary)) +
  geom_boxplot()

nba2021pergame %>%
  ggplot(aes(x=Salary)) +
  geom_histogram()
```

``` {r}
nba2021per36_sub2 <- nba2021per36 %>%
    select(Salary, PTS, AST, TRB)

# Run k-means for k = centers = 3
set.seed(253)
kclust_k3 <- kmeans(nba2021per36_sub2, centers = 5)

# Display the cluter assignments
kclust_k3$cluster


nba2021per36 <- nba2021per36 %>%
    mutate(kclust_3 = factor(kclust_k3$cluster))

ggplot(nba2021per36, aes(x=`Salary`, y= `WS`, color=kclust_3)) +
    geom_point()

set.seed(253)
kclust_k3_scale <- kmeans(scale(nba2021per36_sub2), centers = 5)
nba2021per36 <- nba2021per36 %>%
    mutate(kclust_3_scale = factor(kclust_k3_scale$cluster))

# Visualize the new cluster assignments
ggplot(nba2021per36, aes(x=`Salary`, y= `PTS`, color=kclust_3_scale)) +
    geom_point()

plot3d(nba2021per36$Salary, nba2021per36$PTS, nba2021per36$TRB, col = nba2021per36$kclust_3_scale)
```